Supplemental Figures: Aligning marine species range data to better serve science and conservation

### 2016-09-24 clip_fao/fao_clip_pac.Rmd
added t-test for mean area-per-cells

am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
                                   'v2016/int/am_cells_spp_prob0.csv'))

am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
  dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
  dplyr::select(-genus, -species) %>%
  distinct()

hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
  dplyr::select(loiczid, cellarea, oceanarea)

am_spp_cells2 <- hcaf %>%
  left_join(am_spp_cells, by = 'loiczid')

clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
  filter(lim == 'e_lim') %>%
  dplyr::select(am_sid)

spp_quad_list <- read_csv(file.path(dir_data, 'spp_list_quads.csv'))

spp_area_df <- am_spp_cells2 %>%
  group_by(am_sid) %>%
  summarize(cellarea = sum(cellarea),
            oceanarea = sum(oceanarea))

spp_area_df <- spp_area_df %>%
  inner_join(am_spp_list, by = 'am_sid') %>%
  mutate(area_per_cells = oceanarea/occurcells)

head(spp_area_df)
#                                     am_sid    cellarea   oceanarea reviewed occurcells                   sciname
#                                      <chr>       <dbl>       <dbl>    <chr>      <int>                     <chr>
# 1                                Biv-32126   949988.02   642706.64     null          4        Corbicula japonica
# 2                                Biv-32832  9698932.54  7589672.98     null         23          Polymesoda erosa
# 3                               Bra-235673    16143.75    13895.96        1         10       Dyscritosia secreta
# 4 Chn-013ca97e-46a8-4c5e-b398-8bb4f340f88b 11288439.81 10048833.94     null         76          Lenisquilla lata
# 5 Chn-015453c0-3ff9-477d-ab51-c68d85e65af5 15078672.39 12760166.60     null         13        Clorida albolitura
# 6 Chn-0721ac0c-7d19-46fa-9dcc-02da54310867 35256268.05 33846838.22     null         23 Leiogalathea laevirostris
summary(spp_area_df$oceanarea)
   # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 1806    946600   2921000  10210000   7875000 501300000 
summary(spp_area_df$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.00    6.00   16.00   57.13   49.00 7663.00 
summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    #  190    50540   158000   485300   439100 44700000 
nrow(spp_area_df %>% filter(!str_detect(reviewed, 'null')))
   # 1296 out of 22889 = 5.66%

spp_area_clipped <- spp_area_df %>%
  filter(am_sid %in% clipped_spp$am_sid)
nrow(spp_area_clipped %>% filter(!str_detect(reviewed, 'null')))
   # 22 out of 3208 = .686%

summary(spp_area_clipped$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 134900   4646000   7032000   8960000   9690000 111300000 
summary(spp_area_clipped$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 3.00    6.00   15.00   26.86   33.00  494.00 
summary(spp_area_clipped$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    # 6578   202900   452400   827200   923300 26090000

>>>>

# am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
#                                    'v2016/int/am_cells_spp_prob0.csv')) 
# 
# am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
#   dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
#   mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
#   dplyr::select(-genus, -species) %>%
#   distinct()
# 
# hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
#   dplyr::select(loiczid, cellarea, oceanarea)
# 
# am_spp_cells2 <- hcaf %>%
#   left_join(am_spp_cells, by = 'loiczid')
# 
# 
# spp_area_df <- am_spp_cells2 %>%
#   group_by(am_sid) %>%
#   summarize(cellarea = sum(cellarea),
#             oceanarea = sum(oceanarea))
# 
# spp_area_df <- spp_area_df %>%
#   inner_join(am_spp_list, by = 'am_sid') %>%
#   mutate(area_per_cells = oceanarea/occurcells)
# 
# clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
#   filter(lim == 'e_lim') %>%
#   dplyr::select(am_sid)
# 
# spp_area_df <- spp_area_df %>%
#   mutate(log_area_cells = log(area_per_cells + 1),
#          clipped = am_sid %in% clipped_spp$am_sid)

# write_csv(spp_area_df, 'clip_fao/spp_area_per_cell.csv')

spp_area_df <- read_csv('clip_fao/spp_area_per_cell.csv')

spp_area_clipped <- spp_area_df %>%
  filter(clipped)

### compare ocean areas
summary(spp_area_df$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   #   1806    946600   2921000  10210000   7875000 501300000 
summary(spp_area_clipped$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 134900   4646000   7032000   8960000   9690000 111300000

### compare occurcells
summary(spp_area_df$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.00    6.00   16.00   57.13   49.00 7663.00 
summary(spp_area_clipped$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 3.00    6.00   15.00   26.86   33.00  494.00

### compare area_per_cells
summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    #  190    50540   158000   485300   439100 44700000 
summary(spp_area_clipped$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    # 6578   202900   452400   827200   923300 26090000

### compare reviewed species
nrow(spp_area_df %>% filter(!str_detect(reviewed, 'null')))
   # 1296 out of 22889 = 5.66%
nrow(spp_area_clipped %>% filter(!str_detect(reviewed, 'null')))
   # 22 out of 3208 = .686%

hist(spp_area_clipped$area_per_cells)
hist(spp_area_clipped$log_area_cells)
hist(spp_area_df$log_area_cells)

t.test(x = spp_area_clipped$log_area_cells,
       y = spp_area_df %>% filter(occurcells != 0) %>% .$log_area_cells,
       alternative = 'greater') ### clipped sample mean greater than overall
# t = 49.082, df = 5244.6, p-value < 2.2e-16
# alternative hypothesis: true difference in means is greater than 0
# 95 percent confidence interval:
#  1.058579      Inf
# sample estimates:
# mean of x mean of y 
#  12.97963  11.88434

### 2016-09-24 ms_3_res_disc_concl.Rmd
added t-test for mean area-per-cells

By this measure, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a median predicted range of 452,000 km^2^ per "occurcell" compared to a median predicted range of 158,000 km^2^ per "occurcell" for the overall AquaMaps dataset.  This suggests that the FAO boundaries may not be sufficient to adequately constrain computer-generated ranges.  To reduce the incidence of commission errors due to aggressive extrapolation, it may be desirable to fine-tune the computer model output with additional filters, including ecosystem-based constraints such as Marine Ecoregions of the World [@spalding_marine_2007] or methods such as inverse distance weighting to enforce proximity to known observations.  Expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.

>>>>

By this measure, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a mean predicted range of 827,200 km^2^ per "occurcell" compared to a mean predicted range of 485,300 km^2^ per "occurcell" for the overall AquaMaps dataset.  This suggests that the FAO boundaries may not be sufficient to adequately constrain computer-generated ranges.  To reduce the incidence of commission errors due to aggressive extrapolation, it may be desirable to fine-tune the computer model output with additional filters, such as ecoregional constraints, e.g. Marine Ecoregions of the World [@spalding_marine_2007], or distance-based methods, e.g. inverse distance weighting to enforce proximity to known observations.  Expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.
